### 2020-05-29 paper/main.tex
added model description

@article{He2018ABAE,
title={An Unsupervised Neural Attention Model for Aspect Extraction},
author={Ruidan He, Wee Sun Lee, Hwee Tou Ng, Daniel Dahlmeirer},
year={2017},
}

@article{Mikolov2013W2V,
title={Efficient Estimation of Word Representations in Vector Space},
author={Tomas Mikolov, Kai Chen, Greg Corrado, Jeffrey Dean},
year={2017},
}

@article{Bahdanau2015NMT,
title={Neural Machine Translation by Jointly Learning to Align and Translate},
author={Dzmitry Bahdanau, Kyunghyun Cho, Yoshua Bengio},
year={2015},
}

@article{Hinton2017DRBC,
title={Dynamic Routing Between Capsules},
author={Sara Sabour, Nicholas Frosst, Geoffrey E Hinton},
year={2017},
}

@article{Ganu2009BeyondTheStars,
title={Beyond the Stars: Improving Rating Predictions using Review Text Content},
author={Gayatree Ganu, Noemie Elhadad, Amelie Marian},
year={2009},
booktitle={Proceedings of the 12th International Workshop on the Web and Databases},
}

@article{Zhao2010JMA,
title={Jointly modeling aspects and opinions with a MaxEnt-LDA hybrid},
author={Wayne Xin Zhao, Jing Jiang, Hongfei Yan, Xiaoming Li},
year={2010},
booktitle={Proceedings of the 2010 Conference on Emperical Methods in Natural Language Processing}
}

@article{HuLiu2004,
title={Mining and summarizing customer reviews},
author={Minqing Hu, Bing Liu},
year={2004},
booktitle={Proceedings of the 10th ACM SIGKDD International Conference on Knowledge Dicovery and Data Mining}
}

@article{Zhuang2006,
title={Movie review mining and summarization},
author={Li Zhuang, Feng Jing, Xiao-Yan Zhu},
year={2006},
booktitle={Proceedings of the 15th ACM International Conference on Information and Knowledge Management}
}

@article{Somasundarn2009,
title={Recognizing stances in online debates},
author={Swapna Somasundaran, Janyce Wiebe},
year={2009},
booktitle={Proceedings of the Joint Conference of the 47th Annual Meeting of the ACL and the 4th International Joint Conference on Natural Language Processing of the AFNLP}
}

@article{Qiu2011,
title={Opinion word expansion and target extraction thorugh double propagation},
author={Guang Qiu, Bing Liu, Jiajun Bu, Chun Chen},
year={2011},
journal={Computation Linguistics},
volume={37},
pages={9-27}
}

@article{Jin2009,
title={A novel lexicallized HMM-based learning frameworks for web opinion mining},
author={Wei Jin, Huang Hay Ho},
year={2009},
booktitle={Proceedings of the 26th International Conference on Machine Learning}
}

@article{Li2010,
title={Structure-aware review mining and summarization},
author={Fangtao Li, Chao Han, Minlie Huang, Xiaoyan Zhu, Ying-Ju Xia, Shu Zang, Hao Yu},
year={2010},
booktitle={Proceedings of the 23rd International Conference on Computer Linguistics}
}

@article{Li2010,
title={Structure-aware review mining and summarization},
author={Fangtao Li, Chao Han, Minlie Huang, Xiaoyan Zhu, Ying-Ju Xia, Shu Zang, Hao Yu},
year={2010},
booktitle={Proceedings of the 23rd International Conference on Computer Linguistics}
}

@article{Yin2016,
title={Unsupervised word and dependency path embeddings for aspect term extraction},
author={Yichun Yin, Furu Wei, Li Dong, Kaiming Xu, Ming Zhang Ming Zhou},
year={2016},
booktitle={Proceedings of the 25th International Joint Conference on Artificial Intelligence}
}

@article{Wang2016,
title={Recursive neural conditional random fields for aspect-based sentiment analysis},
author={Wenya Wang, Sinno J. Pan, Daniel Dahlmeier, Xiaokui Xiao},
year={2016},
booktitle={Proceedings of the 2016 Conference on Empirical Methods in Natural Language Processing}
}

@article{Brody2010UAS,
title={An unsupervised aspect-sentiment model for online reviews},
author={Samuel Brody, Noemie Elhadad},
year={2010},
booktitle={Human Language Technologies: The 2010 Annual Conference of the North American Chapter of the Association for Computational Linguistics}
}

@article{Mukherjee2012,
title={Aspect extraction through semi-supervised modelling},
author={Arjun Mukkerjee, Bing Liu},
year={2012},
booktitle={Proceedings of the 50th Annual Meetings of the Association for Computational Linguistics}
}

@article{Chen2014,
title={Aspect extraction with automated prior knowledge learning},
author={Zhiyuan Chen, Arjun Mukherjee, Bing Liu},
year={2014},
booktitle={Proceedings of the 52nd Annual Meeting of the Association for Computational Linguistics}
}

@article{Blei2003,
title={Latent Dirichlet allocation},
author={David Blei, Andrew Ng, Michael Jordan},
year={2003},
journal={Jornal of Machine Learning research},
volume={3},
pages={993-1022}
}

@article{Wang2015,
title={Sentiment-aspect extraction based on restricted Boltzmann machines},
author={Linlin Wang, Kang Liu, Zhu Cao, Jun Zhao, Gerrard de Malo},
year={2015},
booktitle={Proceedings of the 53rd Annual Meeting of the Association for Computational Linguistics and the 7th International Joint Conference on Natural Language Processing}
}

@article{Yan2013,
title={A biterm topic model for short texts},
author={Xiaohui Yan, Jiafeng Guo, Yanyan Lan, Xueqi Cheng},
year={2013},
booktitle={Proceedings of the 22nd International World Wide Web Conference}
}

@article{Mnih2014,
title={Recurrent models of visual attention},
author={Volodymyr Mnih, Nicolas Heess, Alex Graves, Koray Kavukcuoglu},
year={2014},
booktitle={Advanced in Neural information Processing Systems}
}

@article{Rush2015,
title={A neural attention model for sentence summarization},
author={Alexander Rush, Sumit Chopa, Jason Weston},
year={2015},
booktitle={Proceeding of the 2015 Conference on Emperical Methods in Natural Language Processing}
}

@article{Chen2016,
title={Neural sentiment classification with user and product attention},
author={Huimi Chen, Maosong Sun, Cunchao Tu, Yankai Lin, Zhiyuan Liu},
year={2016},
booktitle={Proceeding of the 2016 Conference on Emperical Methods in Natural Language Processing}
}

@article{Herman2015,
title={Teaching machines to read and comperhend},
author={Karl Mortiz Hermann, Tomas Kocisky, Edward Grefenstette, Lasse Espeholt, Will Kay, Mustafa Suleyman, Phil Blunsom},
year={2015},
booktitle={Advanced in Neural Information Processing System}
}

@article{Zhao2018,
title={Investigationg Capsule Network with Dynamic ROuting for Text Classification},
author={Wei Zhao, Jiambo Ye, Min Yang, Zeyang Lei, Suofei Zhang, Zhou Zhao},
year={2018}
}

@article{levenshtein1966binary,
title={Binary Codes Capable of Correcting Deletions, Insertions and Reversals},
author={Levenshtein, Vladimir I.},
year={1966},
month={February},
journal={Soviet Physics Doklady},
volume={10},
pages={709}
}

@article{levenshtein1966dvoichnie,
title={Двоичные коды с исправлением выпадений, вставок и замещений символов},
author={Левенштейн, {\CYRV}ладимир {\CYRI}осифович.},
year={1966},
month={February},
journal={Доклады Академий Наук СССР},
volume={163},
pages={845--848},
number={4}
}

@inproceedings{merity2017pointer,
  title={Pointer sentinel mixture models},
  author={Merity, Stephen and Xiong, Caiming and Bradbury, James and Socher, Richard},
  booktitle={Proceedings of International Conference of Learning Representation},
  year={2017}
}

@inproceedings{habernal2016c4corpus,
  title={C4Corpus: Multilingual Web-size corpus with free license},
  author={Habernal, Ivan and Zayed, Omnia and Gurevych, Iryna},
  booktitle={Proceedings of the Tenth International Conference on Language Resources and Evaluation (LREC'16)},
  pages={914--922},
  year={2016}
}