### 2020-05-30 paper/main.tex
added final paper

\documentclass{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtext}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{breakcites}

\title{Capsule Based Aspect Extraction}
\author{Kirill Krasikov}
\date{May 2020}

\begin{document}
\maketitle
\begin{abstract}
    In the world of data science, more and more data is required to train models. Getting labeled datasets is a very expensive and lengthy process. In this regard, in the problems of topic modeling, aspect extraction models from unstructured text are gaining popularity. These models allow, with minimal effort to annotate aspects to structure the text into a set of topics or intentions. In this paper, a comparison is made of a SotA model based on attention and the model in which the attention mechanism is replaced by a capsule network with dynamic routing.
    Codes for paper: \url{https://github.com/KirillKrasikov/TopicModelingWithCapsNet}.
\end{abstract}

\section{Introduction}
Aspect extraction is an important and challenging task in aspect-based sentiment analysis. For example, in the sentence "The beef was tender and melted in my mouth", the aspect term is "beef". Two sub-tasks are performed in aspects extraction: (1) extracting all aspects terms(e.g., "beef") from a review corpus, (2) clustering aspect terms with similar meaning into categories where each category represents a single aspect(e.g., cluster "beef", "pork", "pasta", and "tomato" into one aspect \textit{food}) \cite{He2018ABAE}. In 2017 year, Unsupervised Aspect Extraction \cite{He2018ABAE} became the dominant unsupervised approach for aspect extraction. In that work word embeddings map words with the same context to nearby points in the embedding space \cite{Mikolov2013W2V}. Then attention mechanism \cite{Bahdanau2015NMT} filter the word embeddings within a sentence and filtered words are used to construct aspect embeddings. The training process for aspect embeddings is analogous with autoencoders, where dimension reduction is used to extract the common factors among embedded sentences and reconstruct each sentence through a linear combination of aspect embeddings. The attention mechanism deemphasizes words that are not part of any aspects, allowing the model to focus on aspect words. That model is called \textit{Attention-based Aspect Extraction} (ABAE). In the same year "Google brain" suggested approach called \textit{Dynamic Routing Between Capsules} \cite{Hinton2017DRBC}. A capsule is a group of neurons whose activity vector represents the instantiation parameters of a specific type of entity such as an object or an object part. The length of the activity vector is used to represent the probability that the entity exists and its orientation is used to represent the instantiation parameters. Active capsules at one level make predictions, via transformation n-dimension arrays, for the instantiation parameters of higher-level capsules. When multiple predictions mutually adjusted, a higher level capsule becomes active. To best performance, iterative routing-by-agreement mechanism is used. A lower-level capsule prefers to send its output to higher level capsules whose activity vectors have a big scalar product with the prediction coming from the lower-level capsule \cite{Hinton2017DRBC}. In this work an ABAE model is implemented using pytorch, then attention mechanism is changed to capsule network and this model was evaluated on Citysearch corpus used by previous works \cite{Ganu2009BeyondTheStars, Brody2010UAS, Zhao2010JMA, He2018ABAE}.

\section{Related Work}
\label{sec:related}
The problem of aspect extraction was well studied in the past decade. Initially, methods were mainly based on manually defined rules. \cite{HuLiu2004} proposed to extract different product features through finding frequent nouns and noun phrases. They also extracted opinion terms by finding the synonyms and antonyms of opinion seed words through WordNet. Following this, a number of methods have been proposed based on frequent item mining and dependency information to extract product aspects \cite{Zhuang2006, Somasundarn2009, Qiu2011}. These models heavily depend on predefined rules which work well only when the aspect terms are restricted to a small group of nouns. Supervised learning approaches generally model aspect extraction as a standard sequence labeling problem. \cite{Jin2009, Li2010} proposed to use hidden Markov models(HMM) and conditional random fields(CRF) respectively with a set of manually-extracted features. More recently, different neural models \cite{Yin2016, Wang2016} were proposed to automatically learn features for CRF-based aspect extraction. Rule-based models are usually not refined enough to categorize the extracted aspect terms. On the other hand, supervised learning requires large amounts of labeled data for training purposes.

Unsupervised approaches, especially topic models, have been proposed subsequently to avoid reliance on labeled data. Generally, the outputs of those models are word distributions or rankings for each aspect. Aspects are naturally obtained without separately performed extraction and categorization. Most existing works \cite{Brody2010UAS, Zhao2010JMA, Mukherjee2012, Chen2014} are based on variants and extensions of LDA \cite{Blei2003}. Recently, \cite{Wang2015} relies on a substantial amount of prior knowledge such as part-of-speech(POS) tagging and sentiment lexicons. Biterm topic models (BTM) that generate co-occurring word pairs was proposed in \cite{Yan2013}.

Attention models \cite{Mnih2014} have recently gained popularity in training neural networks and have been applied to various natural language processing tasks, including machine translation \cite{Bahdanau2015NMT}, sentence summarization \cite{Rush2015}, sentiment classification \cite{Chen2016} and question answering \cite{Herman2015}. Rather than using all available information, attention mechanism aims to focus on the most pertinent information for a task. \cite{He2018ABAE} applies attention to an unsupervised neural model and demonstrate its effectiveness for aspect extraction.

\cite{Zhao2018} explore capsule networks with dynamic routing process \cite{Hinton2017DRBC} for supervised text classification. In this work capsule network with dynamic routing applied to unsupervised neural model and results are compared with others aspect extraction models.

\section{Model Description}
The ultimate goal of Capsule Based Aspect Extraction (CBAE) is to learn a set of aspect embeddings, where each aspect can be interpreted by looking at the nearest words (representative words) in the embedding space. Each word $w$ in vocabulary with a feature vector $e_w \in \R^d$ \cite{He2018ABAE}. Word embeddings are used to map words that often co-occur in a context to points that are close by in the embedding space \cite{Mikolov2013W2V}. The feature vectors associated with the words correspond to the rows of a word embedding matrix $E \in R^{V \times d}$, where $V$ is the vocabulary size. Model tries to learn embeddings of aspects, where aspects share the same embedding space with words. This requires an aspect embedding matrix $T \in R^{K \times d}$, where $K$ - the number of aspects ($K$ is much smaller than $V$). The aspect embeddings are used to approximate the aspect words in the vocabulary, where aspect words are filtered through the capsule network with dynamic routing\footnote{This is the key difference from ABAE \cite{He2018ABAE} , in which the mechanism of attention is used for filtering words.}.
Each input sample to CBAE is a list of indexes for words in a review sentence. First, the model filter away non-aspect words by down-weighting them using a capsule network and then constructs a sentence embedding $z_s$ from weighted word embeddings. Then, the model tries to reconstruct the sentence embedding as a linear combination of aspect embedding from $T$. This process of dimension reduction and reconstruction, where CBAE aims to transform sentence embeddings of the filtered sentences ($z_s$) into their reconstruction ($r_s$) with the least possible amount of distortion, preserves most of the information of the aspect words in the $K$ embedded aspects. This process is described in detail in \cite{He2018ABAE}.

\subsection{Sentence Embedding and Capsule Network with dynamic routing}
Capsule network, depicted in Fig.~\ref{fig:capsule} get sentence embedding vector to input. It consists of four layers: n-gram convolutional layer, primary capsule layer, secondary capsule layer and fully connected capsule layer.

\begin{figure}[!tbh]
    \centering
    \includegraphics[width=1\linewidth]{rect10315.png}
    \caption{The Architecture of capsule network for aspect extraction (CBAE encoder part)}
    \label{fig:capsule}
\end{figure}

N-gram convolutional layer is a standard convolutional layer which extracts n-gram features at different positions of a sentence through various convolutional filters\cite{Zhao2018} and activated by ELU function. Primary capsule layer is the first capsule layer in which the capsules replace the scalar-output feature detectors of CNNs with vector-output capsules to preserve the instantiated parameters such as the local order of words and semantic representation of words. Capsule network tries to address the representational limitation and exponential inefficiencies of convolutions with transformation matrices \cite{Hinton2017DRBC}. It allows the networks to automatically learn child-parent relationships constituting viewpoint invariant knowledge that automatically generalizes to novel viewpoints. Dynamic routing is needed to construct a non-linear map in an iterative manner ensuring that the output of each capsule gets sent to an appropriate parent in the subsequent layer. For each potential parent, the capsule network can increase or decrease the connection strength by dynamic routing, which is more effective than the primitive routing strategies such as max-pooling in CNN that essentially detects whether a feature is present in any position of the text, but loses spatial information about the feature. Secondary capsule layer multiplies transformation matrices to learn child-parent relationships followed by routing agreement to produce parent capsules in the above. Then the capsules from secondary capsule layer gets flattened into fully connected final capsule, which is then activated by \textit{tanh} function and normalized by \textit{softmax}.

\section{Dataset}
The model have been evaluated on \textbf{Citysearch corpus}. This is restaurant review corpus widely used by previous works \cite{Ganu2009BeyondTheStars, Brody2010UAS, Zhao2010JMA}, which contains over 50,000 restaurant reviews from Citysearch New York. \cite{Ganu2009BeyondTheStars} provided a subset of 3,400 sentences from the corpus with manually labeled aspects. These annotated sentences are used for evaluation of aspect identification. There are six manually defined aspect labels: \textit{food}, \textit{staff}, \textit{ambience}, \textit{price}, \textit{anecdotes} and \textit{miscellaneous}. Corpus preprocessing is described in \cite{He2018ABAE}.

\section{Experiments}

\subsection{Metrics}
The model was evaluated by precision, recall and F1 scores. Сlasses of aspects were tagged manually.

\subsection{Experiment Setup}
For the CBAE implementation word embedding matrix E with word vectors trained by word2vec with negative sampling on each dataset. Settings of the word2vec parameters: embedding size = 200, window size = 10 and negative sample size = 5. The aspect embedding matrix initialized with the centroids of clusters results resulting from running $k$-means on word embeddings. Other parameters are initialized randomly. During the training process word embedding matrix $E$ was fixed, other parameters were optimized using \textit{Adam} with learning rate 1e-3 for 10 epochs and batch size of 50. The number of negative samples was set to 20. Following \cite{Brody2010UAS, Zhao2010JMA, He2018ABAE} the number of aspects for the corpus was set to 14. In CBAE like in ABAE, representative  words of an aspect can be found by looking at its nearest words in the embedding space using cosine as the similarity metric. N-gram convolutional layer setting the kernel size to 3, stride to 1 and channels to 200. Primary caps layer has 7 capsules with kernel size = 3, stride = 1 and out channels = 32, Secondary capsule layer has 7 capsules with 32 channels. Dynamic routing performs 3 iterations.

\subsection{Baselines}
To validate the performance of CBAE, the model was compared with next baselines:

1) \textbf{LocLDA} \cite{Brody2010UAS}: This method uses a standard implementation of LDA.

2) \textbf{$k$-means} \cite{He2018ABAE}: Aspect matrix T with k-means centroids of the word embeddings.

3) \textbf{SAS} \cite{Mukherjee2012}: Hybrid topic model that jointly discovers both aspects and aspect-specific opinions.

4) \textbf{BTM} \cite{Yan2013}: Bitermic topic model that is specially designed for short texts such as texts from social media and review sites. The major advantage of BTM over conventional LDA models is that it alleviates the problem of data sparsity in short documents by directly modeling the generation of unordered word-pair co-occurrences (biterms) over the corpus.

5) \textbf{ABAE} \cite{He2018ABAE}: Attention based aspect extraction.

\section{Results}
Tab.~\ref{tab:aspects} presents all 14 aspects inferred by CBAE.

\begin{table}[tbh!]
\begin{center}
\begin{tabular}[t]{l|l}
\hline
Representative Words & Aspects \\
\hline
chicken potato onion tomato sauce           & food          \\
highly grilled seared roasted shrimp        & food          \\
pay money leave didn order                  & price         \\
minute asked manager seated waiter          & staff         \\
wall room space ceiling lit                 & ambience      \\
dish menu entree course flavor              & food          \\
birthday saturday reservation friend friday & anecdotes     \\
year ny experience life nyc                 & miscellaneous \\
service staff waitstaff attentive friendly  & staff         \\
month time minute week hour                 & miscellaneous \\
drink wine sangria nice perfect             & food          \\
gras crust rib pork foie                    & food          \\
chocolate butter fruit sauce sweet          & food          \\
chinese japanese american mexican indian    & food          \\
\hline
\end{tabular}
\caption{List of top representative words for each aspect (left), and aspect labels (right). Aspect labels (right) were assigned manually.}
\label{tab:aspects}
\end{center}
\end{table}

\begin{table}[tbh!]
\begin{center}
\begin{tabular}[t]{l|l|l|l|l}
\hline
 Aspect & Method & Precision & Recall & $F_1$ \\
\hline
          & LocLDA      & 0.90          & 0.65          & 0.75           \\
          & SAS         & 0.88          & 0.77          & 0.82           \\
          & BTM         & 0.93          & 0.75          & 0.82           \\
food      & k-means     & 0.93          & 0.65          & 0.76           \\
          & ABAE        & \textbf{0.95} & 0.74          & 0.83           \\
          & CBAE        & 0.89          & \textbf{0.85} & \textbf{0.87}  \\
\hline
          & LocLDA      & 0.80          & 0.59          & 0.68           \\
          & SAS         & 0.77          & 0.56          & 0.65           \\
          & BTM         & \textbf{0.83} & 0.58          & 0.68           \\
staff     & k-means     & 0.79          & 0.69          & 0.66           \\
          & ABAE        & 0.80          & 0.73          & \textbf{0.77}  \\
          & CBAE        & 0.68          & \textbf{0.75} & 0.71           \\
\hline
          & LocLDA      & 0.60          & 0.68          & 0.64           \\
          & SAS         & 0.78          & 0.54          & 0.65           \\
          & BTM         & 0.81          & 0.6           & 0.69           \\
ambience  & k-means     & 0.73          & 0.64          & 0.68           \\
          & ABAE        & 0.82          & \textbf{0.70} & \textbf{0.74}  \\
          & CBAE        & \textbf{0.83} & 0.62          & 0.71           \\
\hline
\end{tabular}
\caption{Aspect identification results. The result of LocLDA is taken from \cite{Zhao2010JMA}; the result of SAS is taken from \cite{Wang2015}; the results of BTM, k-means and ABAE are taken from \cite{He2018ABAE}}.
\label{tab:eval}
\end{center}
\end{table}

Evaluation criterion of sentence-level-aspects identification is to judge how well the predictions match the true labels. It is measured by precision, recall and $F_1$ scores. The results are shown in Tab .~\ref{tab:eval}

\section{Conclusion}
The analysis of the results shows that the attention-based model in general has more scores.
Greater recall of capsule-based model is determined by the lower severity of aspects relative to the attention-based model (for example \textit{nice} and \textit{perfect} with \textit{drink} and \textit{wine} or \textit{leave} with \textit{pay} and \textit{money}). Perhaps the orchestration of these models will make it possible to obtain higher marks in the problem of aspect extraction. \\ \\ \\ \\ \\ \\

\bibliographystyle{apalike}
\bibliography{lit}
\end{document}